---
title: "A World in Motion"
subtitle: "Global Forced Displacement Flows, 2000–2025"
author: "Soe Min Thein (Vincent)"
date: "April 20, 2026"
format:
  beamer:
    theme: Madrid
    colortheme: seahorse
    aspectratio: 169
    header-includes: |
      \setbeamertemplate{footline}[frame number]
      \setbeamerfont{title}{size=\large}
      \setbeamerfont{subtitle}{size=\normalsize}
      \hypersetup{colorlinks=true, urlcolor=blue}
      
knitr:
  opts_chunk:
    echo: false
    eval: true
    prompt: false
    tidy: false
    message: false
    warning: false
    fig.align: 'center'
---

```{r}
# Load necessary libraries
library(tidyverse)
library(readxl)
library(knitr)
library(patchwork)
```

```{r}
# Read the data from the Excel file
data <- read_excel("data/UNHCR_Flow_Data.xlsx", sheet = "DATA")

# Data cleaning and preparation
# Filter the data for REF + ASY combined, and years 2000-2025
filtered_data <- data |> 
  filter(PT %in% c("REF", "ASY"), Year >= 2000, Year <= 2025)

# Custom theme 
mytheme <- theme_minimal() +
  theme(
    plot.title = element_text(size = 15, face = "bold"),
    plot.subtitle = element_text(size = 10),
    plot.caption = element_text(size = 8, hjust = 0),
    axis.text = element_text(size = 9),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    panel.grid.minor = element_blank()
  )
```


## Overview

**Research Question**

*Where do forcibly displaced people come from, and where do they go?*

**Findings**

- Global displacement flows have grown dramatically over the past 25 years, driven by a series of major crises
- A small number of countries account for the vast majority of all outflows
- Displaced people overwhelmingly seek refuge in **neighboring countries**, not wealthy Western nations – though the US and Germany rank high due to large asylum application volumes
- The biggest displacement corridors reflect geography, not global burden-sharing


## Data

**Source:** [UNHCR Forced Displacement Flow Dataset](https://www.unhcr.org/refugee-statistics/insights/explainers/forcibly-displaced-flow-data.html)

**Key Variables**

- `OriginName` – country people fled *from*
- `AsylumName` – country people fled *to*
- `Count` – this number of people in that flow
- `Year` – year of flow

**Note**

- Coverage – 2000-2025 (full dataset includes 1962-2025)
- Observations – 86,275 rows (after filtering for REF + ASY and years 2000-2025)
- Population types – Refugees (REF) + Asylum-seekers (ASY), the two largest 
  and most consistently reported categories

*Pre-2000 data excluded due to sparse and inconsistent coverage in many countries.*


## A World in Motion

```{r}
#| fig-width: 10
#| fig-height: 3.5
#| out-width: "100%"

# Visual 1 ----------------------------------------------------------------
# Group by Year and summarize total Count
yearly_trends <- filtered_data |> 
  group_by(Year) |> 
  summarize(Total_Count = sum(Count, na.rm = TRUE))

# Add crisis labels data
crisis_labels <- data.frame(
  Year  = c(2013, 2017, 2022),
  Label = c("Syria crisis", "Myanmar\n(Rohingya crisis)", "Ukraine war"),
  y     = c(5.0, 5.5, 7.5)  # adjust these numbers after you see the plot
)

ggplot(yearly_trends, aes(x = Year, y = Total_Count / 1000000)) +
  # shaded area under the line
  geom_area(fill = "steelblue", alpha = 0.15) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 2) +
  # crisis annotations
  geom_vline(xintercept = c(2013, 2017, 2022), 
             linetype = "dashed", color = "gray60", linewidth = 0.5) +
  geom_text(data = crisis_labels,
            aes(x = Year, y = y, label = Label),
            hjust = -0.1, size = 2.8, color = "gray40") +
  scale_x_continuous(
    limits = c(2000, 2025),
    breaks = seq(2000, 2025, by = 5),
    expand = expansion(mult = c(0, 0.05))
    ) +
  scale_y_continuous(
    limits = c(0, 9), 
    breaks = seq(0, 8, by = 2),
    expand = expansion(mult = c(0, 0.01))
    ) +
  labs(title = "Trend of Total Flows by Year",
       subtitle = "Global forced displacement flows, 2000–2025 (in millions)",
       caption = "Source: UNHCR Forced displacement flow dataset",
       x = "Year",
       y = "People (Millions)") +
  mytheme
```
Displacement is not random, it spikes sharply around identifiable crises and has never returned to pre-crisis baselines. 

## Where Are People Fleeing From and Going To?

```{r}
#| fig-width: 13
#| fig-height: 4
#| out-width: "100%"

# Visual 2 ----------------------------------------------------------------
# Step 1: identify top 10 unique origin countries by total combined count
top10_origins <- filtered_data |> 
  group_by(OriginName) |> 
  summarize(Total_Count = sum(Count, na.rm = TRUE), .groups = "drop") |> 
  arrange(desc(Total_Count)) |> 
  slice_head(n = 10) |> 
  pull(OriginName)  # extract just the country names as a vector

# Step 2: filter filtered_data to those 10 countries, then group by PT
top_origin_countries <- filtered_data |> 
  filter(OriginName %in% top10_origins) |>
  group_by(OriginName, PT) |> 
  summarize(Total_Count = sum(Count, na.rm = TRUE), .groups = "drop") |>
  group_by(OriginName) |>
  mutate(Country_Total = sum(Total_Count)) |>  # combined REF + ASY per country
  ungroup()

# Plot
p2 <- ggplot(top_origin_countries, 
       aes(x = Total_Count / 1000000,
           y = reorder(OriginName, Country_Total),
           fill = PT)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = c("REF" = "aquamarine4", "ASY" = "aquamarine2"),
                    labels = c("REF" = "Refugee", "ASY" = "Asylum-seeker")) +
  scale_x_continuous(
    limits = c(0, 10), 
    breaks = seq(0, 10, by = 2),
    expand = expansion(mult = c(0, 0.01))
    ) +
  labs(title = "Where Are People Fleeing From?",
       subtitle = "Top 10 origin countries by cumulative outflows, 2000–2025",
       caption = "Source: UNHCR Forced displacement flow dataset",
       x = "People (Millions)", y = NULL,
       fill = NULL) +
  mytheme

# Visual 3 ----------------------------------------------------------------
# Step 1: identify top 10 unique asylum countries by total combined count
top10_asylum <- filtered_data |> 
  group_by(AsylumName) |> 
  summarize(Total_Count = sum(Count, na.rm = TRUE), .groups = "drop") |> 
  arrange(desc(Total_Count)) |> 
  slice_head(n = 10) |> 
  pull(AsylumName)  # extract just the country names as a vector

# Step 2: filter filtered_data to those 10 countries, then group by PT
top_asylum_countries <- filtered_data |> 
  filter(AsylumName %in% top10_asylum) |>
  group_by(AsylumName, PT) |> 
  summarize(Total_Count = sum(Count, na.rm = TRUE), .groups = "drop") |>
  group_by(AsylumName) |>
  mutate(Country_Total = sum(Total_Count)) |>  # combined REF + ASY per country
  ungroup()

# Plot
p3 <- ggplot(top_asylum_countries, 
       aes(x = Total_Count / 1000000, 
           y = reorder(AsylumName, Country_Total), 
           fill = PT)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = c("REF" = "lightsteelblue4", "ASY" = "lightsteelblue2"),
                    labels = c("REF" = "Refugee", "ASY" = "Asylum-seeker")) +
  scale_x_continuous(
    limits = c(0, 6), 
    breaks = seq(0, 6, by = 2),
    expand = expansion(mult = c(0, 0.01))
  ) +
  labs(title = "Where Are People Going?",
       subtitle = "Top 10 asylum countries by cumulative inflows, 2000–2025",
       caption = "Source: UNHCR Forced displacement flow dataset",
       x = "People (Millions)", y = NULL,
       fill = NULL) +
  mytheme

p2 + p3
```

Displaced people overwhelmingly seek refuge in **neighboring countries** – though wealthy nations absorb significant asylum applications.

Syria and Ukraine dominate outflows since 2000. 
On the receiving end, Türkiye and Uganda reflect geographic proximity, 
while the US and Germany reflect high-volume asylum processing systems.

## Displacement Follows Borders
:::: {.columns}

::: {.column width = "20%"}
The biggest displacement corridors reveal a consistent pattern that **people flee to the nearest safe country**.
:::
  
::: {.column width = "80%"}

```{r}
top_origin_asylum_pairs <- filtered_data |> 
  group_by(OriginName, AsylumName) |> 
  summarize(Total_Count = sum(Count, na.rm = TRUE), .groups = "drop") |> 
  arrange(desc(Total_Count)) |> 
  slice_head(n = 10) |>
  mutate(Rank = row_number(),
         Total_Count = scales::comma(Total_Count)) |>
  select(Rank, everything())

kable(top_origin_asylum_pairs,
      col.names = c("Rank", "Origin Country", "Asylum Country", "Total Displaced"),
      align = c("c", "l", "l", "r"),
      caption = "Top 10 Origin–Asylum Corridors, Cumulative 2000–2025")
```
:::

::::


## Conclusion

**What did we find?**

- Global forced displacement flows have grown sharply since 2000, with major spikes around Syria (2013–2016), Myanmar (2017), and Ukraine (2022)
- Syria and Ukraine generate the largest outflows – together accounting for nearly half of all displacement flows since 2000
- The countries absorbing the most displaced people are **neighboring low and middle-income countries** – with the US and Germany as exceptions driven by high asylum application volumes
- Top displacement corridors are defined by **geography and proximity**, not international burden-sharing agreements

**What does it mean?**

Global displacement is a structural, ongoing challenge – not a series of isolated emergencies. The mismatch between where displacement originates and where international attention focuses has real implications for humanitarian resource allocation and policy.

